Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2015-02-21 papers/wdmerger_I/paper.tex
Our old strategy for tracking the stellar centers of mass failed when the stars got significantly far away from their expected circular orbit locations. This has been fixed by storing the center of mass at every time step and updating it with the old COM velocity. We then refine our guess using the location-weighted sums over the domain. We also make the cutoff more physically meaningful by using the approximation to the Roche radius of Eggleton (1983). To ensure continuity at a restart, I restored Mike's old solution of writing out the mass and COM to a file in the checkpoint, and reading it

We track a number of diagnostic quantities at the end of every coarse grid timestep. 
For all simulations, we record the total energy (including the breakdown into
its components: kinetic, internal, gravitational potential, rotation; we note
that for the diagnostics we actually use $(\rho E)$ for calculation of the total energy,
rather than explicitly calculating the sum of kinetic and internal, as this is
the quantity that should be explicitly conserved), 
the total angular momentum, and the center of mass of the system. 
For stellar calculations, we separately record diagnostic 
information about the stars, including their centers of mass, 
and the volume contained within each star.

>>>>

We track a number of diagnostic quantities at the end of every coarse grid timestep. 
For all simulations, we record the total energy (including the breakdown into
its components: kinetic, internal, gravitational potential, rotation; we note
that for the diagnostics we actually use $(\rho E)$ for calculation of the total energy,
rather than explicitly calculating the sum of kinetic and internal, as this is
the quantity that should be explicitly conserved), 
the total angular momentum, and the center of mass of the system. 
For stellar calculations, we separately record diagnostic 
information about the stars. Our strategy for tracking their 
locations is as follows: at the beginning of the calculation, we store the 
physical center of mass $\mathbf{x}_{\text{COM}}$ of the stars as determined 
by Kepler's third law. We also store the space velocity $\mathbf{v}_{\text{COM}}$ 
of the stars. Then, at each new time step we make a preliminary guess for their 
location by updating the location using the old velocity, 
$\mathbf{x}_{\text{COM}} \rightarrow \mathbf{x}_{\text{COM}} + \mathbf{v}_{\text{COM}} \Delta t$.
We then refine our guess for the location and velocity of each star by computing a
location-weighted sum of the mass and velocity over the computational domain. 
To do this, we need a cutoff for determining what counts as part of the primary 
and what counts as part of the secondary. For this we can use the formula provided by 
\citet{eggleton:1983} for the effective Roche radius $r_L$ of a star in a binary,
\begin{equation}
  \frac{r_L}{a} = \frac{0.49 q^{2/3}}{0.6^{2/3} + \text{ln}(1 + q^{1/3})}.
\end{equation}
In this formula we can use $q = M_S / M_P$ for obtaining the Roche radius of the secondary,
and use the inverse value of $q$ to obtain the Roche radius of the primary.
To a reasonable approximation, all material within the effective Roche radius
of each star can be counted as belonging to that star. From this we obtain the mass
of each star as well as its location and velocity.
